---
title: Setup data for analysis of the GSA Auctions experiment^[`r paste("<",system('grep -n github .git/config | cut -d "=" -f 2',intern=TRUE),">",sep="")`]
author: Jake
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_document:
    highlight: pygments
    keep_md: true
---

```{r include=FALSE, cache=FALSE}
# Some customization.  You can alter or delete as desired (if you know what you are doing).
# knitr settings to control how R chunks work.

## To make the html file do
## render("datasetup.Rmd",output_format=html_document(fig_retina=FALSE))
## To make the pdf file do
## render("datasetup.Rmd",output_format=pdf_document())

require(knitr)
opts_chunk$set(
  tidy=FALSE,     # display code as typed
  size="small",    # slightly smaller font for code
  echo=TRUE,
  results='markup',
  strip.white=TRUE,
  fig.path='figs/fig',
  cache=FALSE,
  highlight=TRUE,
  width.cutoff=132,
  size='footnotesize',
  out.width='.9\\textwidth',
  message=FALSE,
  comment=NA)
```

# Design

For each item posted for auction with no previous bids, an algorithm identifies a population of
potential bidders (people who had previously used the system who had bid on an
item of the same type in the past 3 months). Within this group, half are assigned to receive an email informing them that said item's auction is ending soon. The other half receives no email.

# Outcomes

The key outcome for now is whether or not a person makes a bid.

# Pre-processing of files

I don't think that I found any semi-colons all alone in the files that matter
to us here.^[These next bits of bash script are not run here because they only need
to be run once per set of files. Probably these lines belong in a
pre-processing file, maybe just a bash script.]

```{r eval=FALSE, engine='bash'}
cd Data
egrep  '[^;];[^;]' EMAILQ\ _1101.TXT
egrep  '[^;];[^;]' BIDS_1101.TXT
```

So, this means that I can just read them into R with a single ';' as a separator and then delete empty columns. It might eventually be faster to pre-process the files with 'sed' or 'python' or something.

It turns out I was having trouble reading in BIDS_1101.TXT (one of the lines had some seemingly hidden characters that was causing fread to bomb), so I did a little cleanup of that file in the process of diagnosing the problem.

First, replace all ';;;' or ';;;;' with ';' (since I had already verified that we never had a ';' by itself). I discovered one weird case with 4 ';' 

```{r eval=FALSE, engine='bash'}
egrep  '[^;;;;];;;;[^;;;;]' *.TXT # first just check for four in a row
sed -i.bak1  's/;;;;/;/g' BIDS_1101.TXT
sed -i.bak2  's/;;;/;/g' BIDS_1101.TXT
```

Then, remove excess whitespace at end of line (this is just to make the process of looking for the problem easier) and excess dots

```{r eval=FALSE,engine='bash'}
sed -i.bak3 's/ *$//' BIDS_1101.TXT
sed -i.bak4 's/\.\.\./\./' BIDS_1101.TXT
```



# Reading the files into R

```{r}
## Specific file names need to be changed by hand when the files change unless we automate this some how.
library(data.table,quietly=TRUE) ## using data.table for speed with large datasets

emailq<-fread("Data/EMAILQ\ _1101.TXT",sep=";",header=TRUE,colClasses="character")
## emailq<-emailq[,which(unlist(lapply(emailq, function(x){!all(is.na(x))}))),with=F]
emailq<-emailq[,which(unlist(lapply(emailq, function(x){!all(x=="")}))),with=F]
setnames(emailq,names(emailq),make.names(gsub("\\.","",names(emailq)))) ## make nicer names
setnames(emailq,"REG.SALE.LOT..","RegSaleLot")
str(emailq)
```


We can see that the experimental pool varied a lot across the different items including some items for which only one email was eligible for sending. In those cases, the only email eligible was sent except for a case with SENT=="E" (I'm not sure what E means). So, we will have to remove these rows from consideration since there is no comparison group.

```{r}

regSaleLotsTab<-table(emailq[,RegSaleLot])
sort(regSaleLotsTab,decreasing=TRUE)[1:10]
sort(table(emailq[,RegSaleLot]))[1:10]

## SENT is Control:
table(emailq[SENT.DTE=="00000000",SENT])

emailq[RegSaleLot %in% names(regSaleLotsTab[regSaleLotsTab==1]),list(RegSaleLot,SENT)]

emailq <- emailq[!(RegSaleLot %in% names(regSaleLotsTab[regSaleLotsTab==1])),]
stopifnot(any(table(emailq[,RegSaleLot])!=1))

```

Notice that the same person can be in the experiment multiple times (either as
treatment or control). This means that the causal estimand will be a bit
difficult to define and also that statistical inference will not be
straightforward.

```{r}
table(table(emailq$USER.ID))
```

There is a BID column in 'emailq' but I don't know what it is.

In the BIDS dataset, we have a column for REG, for SALE.NUMBER and LOT. We want a `r unique(nchar(emailq[,RegSaleLot]))` character code from the bids data so that we can combine information about the people making bids and the bids themselves (whether or not a control or treatment assigned email address bid on an item).

```{r}
bids<-fread("Data/BIDS_1101.TXT",sep=";",header=TRUE,colClasses="character")
setnames(bids,names(bids),make.names(gsub("\\.","",names(bids)))) ## make nicer names
str(bids)

## Testing on a small piece of the dataset
set.seed(12345)
blah<-bids[sample(1:nrow(bids),10),]
blah[,RegSaleLot:=paste(REG,SALE.NUMBER,LOT,sep=""),by=1:nrow(blah)]
stopifnot(unique(nchar(blah[,RegSaleLot]))==15)

bids[,RegSaleLot:=paste(REG,SALE.NUMBER,LOT,sep=""),by=1:nrow(bids)]
stopifnot(unique(nchar(bids[,RegSaleLot]))==15)

```

Is there a user id from the email file that exists in the bids file showing that we have bid for a given sale?

First, make a unique id for user and sale:

```{r}

bids[,ID:=paste(RegSaleLot,USER.ID,sep=""),by=1:nrow(bids)]
emailq[,ID:=paste(RegSaleLot,USER.ID,sep=""),by=1:nrow(emailq)]
stopifnot(all(table(emailq$ID)==1)) ## good, uniquely identifies rows of emailq
```

Then collapse bid to the sale/user level (sometimes people gave many bids for the same sale):

```{r}
bids[,bidamt:=as.numeric(BID.AMT)]
bidsItemUser<-bids[,mean(bidamt),by=ID]
setnames(bidsItemUser,"V1","meanbidamt")
stopifnot(all(table(bidsItemUser$ID)==1))
```

Then add the mean bid amt to the data recording whether an email was sent:

```{r}
setkey(bidsItemUser,ID)
setkey(emailq,ID)

## Merge keeping the rows of emailq
mergedat<-bidsItemUser[emailq]
stopifnot(all(table(mergedat$ID)==1))
str(mergedat)
summary(mergedat$meanbidamt)
table(!is.na(mergedat$meanbidamt),mergedat$SENT,exclude=c())
```

Save the data for analysis:

```{r}
write.csv(mergedat,file="Data/mergedat.csv")
save(mergedat,file="Data/mergedat.rda")
```





